Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-30 PAPER.tex
made more of Jonnie's edits, namely removed text referring to 'strong' vs. 'weak' modes

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss3.pdf}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss4.pdf}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- 
foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of foregrounds peaks at a $k$-mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare empirically estimated inverse covariance weighting 
where $\textbf{C}$ is derived from the data (left), and projecting out the first eigenmode only (right). In the former 
case, signal loss arises from the coupling of the eigenmodes of $\widehat{\textbf{C}}$ to the data. For an empirically estimated $\widehat{\textbf{C}}$, its eigenvalues differ from those of the true covariance, where the weakest (EoR-dominated) eigenmodes are the most strongly coupled to the data and can lead to the most loss.
Hence, there is negligible signal loss when assigning identical weights to all eigenmodes except the first, since we are not using the relative weights of the weaker eigenmodes.}
	\label{fig:toy_sigloss3}
\end{figure*}

@@

Because $\widehat{\textbf{C}}$ is estimated from the data, its eigenvectors and eigenvalues are strongly coupled to the particular data realization that was used to compute it. For example, the strongest mode of $\widehat{\textbf{C}}$ (highest eigenvalue) describes the sinusoid foreground mode in the toy model (the peak in Figure \ref{fig:toy_sigloss3}). In Figure \ref{fig:toy_sigloss12} we show the estimated covariances of our toy model datasets along with $\widehat{\textbf{C}}^{-1}$ weighted data. The foreground sinusoid is clearly visible in $\widehat{\textbf{C}}_{\rm FG}$.

In general, the strongest eigenmodes of $\widehat{\textbf{C}}$ typically describe bright foregrounds --- the most prominent shapes in a dataset. For these ``strong" modes, where foregrounds outshine the EoR signal, down-weighting is beneficial. In some sense, we desire signal loss in this regime, if by `signal' we mean `foregrounds.' In this case it is beneficial for the ``strong" eigenmodes to be coupled to the data in a way that produces loss. For our toy model, the successful suppression of the foreground mode is demonstrated in Figure \ref{fig:toy_sigloss3} by the missing foreground peak in the weighted power spectrum estimate (green).

@@

The danger of an empirically estimated covariance matrix comes mostly from not being able to describe the ``weak" eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the coupling between these modes to the data realization leads to the overfitting and subtraction of the EoR signal. More specifically, the coupling between the estimated covariance and the data is anti-correlated in nature (which is explained in more detail in Section \ref{sec:siglossmethod}), which leads to loss. Mis-estimating $\textbf{C}$ for EoR-dominated eigenmodes is therefore more harmful than for FG-dominated modes, and since the ``weak" modes of an eigenspectrum are typically EoR-dominated, using this part of the spectrum for weighting is most dangerous.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield minimal 
signal loss. Recall that our toy model foreground is a sinusoid, so it can be perfectly described by a single eigenmode. Using the full dataset's (foreground plus EoR signal) empirical covariance, we can 
project out the first eigenmode and then flatten the rest of the spectrum to have eigenvalues of 1, thereby down-weighting the foreground-dominated mode more than the rest of the modes. Hence, we are changing the weaker part of the spectrum to be less coupled to the data, limiting the amount of over-fitting that can happen for those modes (i.e. only allowing over-fitting to occur for the first mode).

@@

We have shown how signal loss can arise due to the coupling of weak eigenmodes (EoR-dominated modes) to the data. We will next show how this effect is exacerbated by reducing the total number of independent 
samples in a dataset.

@@

The resulting eigenspectrum as compared to the green curve (FG + EoR) in Figure \ref{fig:toy_sigloss2} is shown in Figure 
\ref{fig:toy_sigloss15}. With 
fringe-rate filtering resulting in fewer independent modes, it becomes more difficult for the empirical covariance to estimate the 
true covariance matrix of the fringe-rate filtered data. This can be quantified by evaluating a convergence metric $\varepsilon(\Chat)$ for the empirical covariance, which we define as
\begin{equation}
\label{eq:converge}
\varepsilon (\Chat) \equiv \sqrt{\frac{\sum_{ij} (\widehat{C}_{ij} - {C}_{ij})^2}{\sum_{ij} {C}_{ij}^2}},
\end{equation}
where $\C$ is the true covariance matrix. In Figure \ref{fig:toy_sigloss16} we show this convergence statistic as a function of the number of independent ensemble realizations in one's simulations (horizontal axis) and the number of independent samples in the data following fringe-rate filtering (different colors). With more independent time samples (i.e. more realizations) in the data, one converges to the true fringe-rate filtered covariance more quickly.

@@

Just as important as the eigenvalues are the eigenvectors of our empirical covariances. In general, different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix whose true form is a diagonal matrix with eigenvalues spanning four orders of magnitude. We use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$ for the empirical eigenvectors $\widehat{\textbf{v}}$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves  in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. One sees that the stronger eigenmodes converge to their true eigenvectors more quickly. With only a small number of realizations, these empirically estimated modes already retain little correlation with the specific realizations of data that were used to form the empirical covariance. As we will see in the next section, using only the strongest eigenmodes, which are less coupled to data realizations, minimizes signal loss. In contrast, the weaker eigenmodes retain more memory of the data realizations, which leads to correlations that induce signal loss. Said differently, a steep covariance eigenspectrum can be especially dangerous because it is the ``weak" modes that are both EoR-dominated and converge the slowest and are therefore susceptible to the most loss.  %\acl{Think about whether we need to do the Fourier space version as well.}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.31\textwidth]{plots/toy_sigloss15.pdf}
	\caption{Eigenspectrum of $\widehat{\textbf{C}}_{\rm FG+EoR}$, in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green). The dashed, steep eigenspectrum has a greater risk of signal loss because its weak eigenmodes are more strongly coupled to the data than those of the solid eigenspectrum.}
	\label{fig:toy_sigloss15}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/toy_sigloss17.pdf}
	\caption{The convergence level, as defined in Equation \eqref{eq:converge_eig}, of empirically estimated eigenvectors for different numbers of mock data realizations. The colors span from the 0th eigenmode (has the highest eigenvalue) to the 19th eigenmode (has the lowest eigenvalue), where they are ordered by eigenvalue in descending order. This figure shows that ``strong" eigenmodes (those with the highest eigenvalues) converge more quickly than ``weak" eigenmodes, implying that weighting by empirically estimated ``weak" modes poses the most risk for signal loss.}
	\label{fig:toy_sigloss17}
\end{figure}

@@

The third panel (bottom left) in Figure \ref{fig:toy_sigloss8} minimizes signal loss a different way --- 
by only using the first three eigenmodes of the estimated covariance. Recalling that our toy model foregrounds can be described entirely by the first eigenmode, this 
method intentionally projects out the strongest modes only by replacing all but the three highest weights in the 
eigenspectrum with $1$'s (equal weights). Again, avoiding the over-fitting of ``weak" modes which are coupled to the data results in negligible signal loss. However, we do 
not perfectly recover the shape of the EoR power spectrum because we lost information when ignoring the relative weights of most modes. While this case is illuminating for the toy model, in practice it is not obvious which eigenmodes are FG or EoR dominated (and they could be mixed as well), so determining which subset of modes to down-weight is not trivial.

The last regularization scheme we are highlighting here is setting $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} \circ \textbf{I}$ (element-wise multiplication), or inverse variance weighting (keeping only the diagonal elements of $\widehat{\textbf{C}}$). In the bottom right 
panel of Figure \ref{fig:toy_sigloss8}, we see that this method does not down-weight the foregrounds --- this regularization altered $\widehat{\textbf{C}}$ in a way where it is no longer coupled to either the ``strong" or ``weak" eigenmodes. For this toy model, 
our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by 
the identity, element-wise, results in a diagonal matrix, meaning we do not have any correlation information. Because of this, we do a poor job 
suppressing the foreground. But because we de-coupled the whole eigenspectrum from the data, we also avoid signal loss. Although this method did not successfully recover the EoR signal for this particular simulation, it is important that we show that there 
are many options for estimating a covariance matrix, and some may down-weight certain eigenmodes more than others based on the spectral nature 
of the components in a dataset. 
%One may imagine a situation where a particular systematic is contained to an isolated 
%frequency (such as radio frequency interference or crosstalk). In such a case, preserving only the diagonal elements of $
%\widehat{\textbf{C}}$ would be an effective way of removing this contamination.

@@

{\bf Small injection:}
%In this regime, the first term in Equation \eqref{eq:crossterm} is very similar to the power spectrum of data alone, namely $\text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \sim \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}$. They differ only in their covariance matrices, and when the injected signal is very small compared to the data, their difference can be neglected. 
%This is shown by the black and gray curves converging in both panels of Figure \ref{fig:sigloss_terms}. 
In this regime, the cross-terms (red) behave as noise averaged over a finite number of samples. Output values are Gaussian distributed around zero, spanning a range of values set by the injection level. This is because $\widehat{\textbf{R}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below). In fact, for the uniformly weighted case, the cross-term  $\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}$ is well modeled as a symmetric distribution with zero mean and width $\sqrt{P_e}\sqrt{P_x}$. We also note that in this regime, $\widehat{\textbf{P}}_{r}$ (black) approaches the data level (gray) as expected. %In this limit we might imagine that the cross-correlation should also be small, enabling us to recover just the EoR term. However, examining the cross-term plotted in Figure \ref{fig:sigloss_terms} (red) in the identity case (right plot), we see that this term is still large enough to have a significant impact on $P_{\rm out}$. It is in this regime where we're dominated by biases from the cross-terms, causing the difference between the signal-only term (green) and $P_{\rm out}$ (black). We also note that the cross-terms behave as noise at small injections even in the inverse covariance weighted case (left plot), spanning a range of values set by the injection level. This is because $\widehat{\textbf{C}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below).

@@

{\bf In between:}
When the injected signal is of a similar amplitude to the data by itself, the situation becomes less straightforward. We see that 
the weighted injected power spectrum component mirrors the input power indicating little loss (i.e. the green curve follows the dotted black line), eventually 
departing from unity when the injected amplitude is well above the level of the data power spectrum. However, 
in this regime the cross-term has nearly the same amplitude, but with a negative sign. As explained below, this negativity is the result of cross-correlating inverse covariance weighted terms.  This negative component drives down the $P_{\rm out}$ estimator (black). We note that in \citetalias{ali_et_al2015}, signal loss was computed by only looking at the second term in Equation \eqref{eq:crossterm} (green), which incorrectly implies no loss at the data level. Ignoring the effect of the negative power from the cross-terms is the main reason for under-estimating power spectrum limits in \citetalias{ali_et_al2015}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.pdf}
	\caption{Illustration of the power spectrum amplitude of various power spectrum terms as a function of injected EoR power level summed into the 
data. The net output ($P_{\rm out}=\widehat{\textbf{P}}_r$) is shown in black, the individual terms, as defined by Equation 
\eqref{eq:crossterm},  are shown in blue, red, green; the dotted black line indicates 1:1 input to output mapping.  The 
empirically estimated inverse covariance weighted case used in \citetalias{ali_et_al2015} is on the left and the uniform-weighted 
case is on the right. The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we fit smooth polynomials to our data points to make an illustrative example.   The gray 
horizontal line is the power spectrum value of data alone (doesn't depend on injected power). The green signal-signal 
component is the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than the net output because the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlations between injection and data are not dominant, the cross-terms have a noise-like 
term with width $\sqrt{P_e}\sqrt{P_x}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to the couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data level) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss at the data level for 
the weighted case). }
	\label{fig:sigloss_terms}
\end{figure*}

@@

\noindent where we can neglect the first term because $\textbf{x}$ is small.  Without loss of generality, we will assume
an eigenbasis of $\textbf{e}$, so that $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$ is diagonal. The middle 
two terms, however, can have power in their off-diagonal terms due to the fact that, when averaging over a finite
ensemble, $\langle\textbf{xe}^\dagger\rangle_t$ is not zero.  As shown in Appendix C of \citet{parsons_et_al2014}%\footnote{This same appendix also gives a now-prescient warning about signal loss and the dangers of noise and sample variance in inverse covariance.}
, to leading order the inversion of a diagonal-dominant matrix like $\widehat{\textbf{C}}_{r}$ (from $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$) with smaller
off-diagonal terms results in a new diagonal-dominant matrix with negative off-diagonal terms. These off-diagonal
terms depend on both $\textbf{x}$ and $\textbf{e}$. Then, when $\widehat{\textbf{C}}^{-1}_{r}$ is multiplied into $\textbf{x}$,
the result is a vector that is similar to $\textbf{x}$ but
contains a residual correlation to $\textbf{e}$ from the off-diagonal components of $\widehat{\textbf{C}}^{-1}_{r}$. The
correlation is negative because the product $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ effectively squares the $\textbf{x}$-dependence
of the off-diagonal terms in $\widehat{\textbf{C}}^{-1}_{r}$ while retaining the negative sign that arose from the inversion
of a diagonal-dominant matrix.
% XXX [ARP: we may want to write this all out explicitly in an appendix]

@@

The implementation details of the injection process require some more detailed explanation. In our code, we add a new realization of EoR to each independent bootstrap of data with the goal of simultaneously capturing variance due to noise and signal loss. To limit computing time we perform $20$ realizations of each $P_{\rm in}$ level. We also run $50$ total EoR injection levels, yielding $P_{\rm in}$ values that range from $\sim$$10^{5}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$ to $\sim$10$^{11}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$, resulting in a total of $1000$ data points on our $P_{\rm in}$ vs. $P_{\rm out}$ grid.

@@

Power spectrum results for the noise simulation, which uses our full power spectrum pipeline, are shown in Figure 
\ref{fig:ps_noise}, where the black and gray points represent pre-signal loss estimated positive and negative power spectrum values, respectively (with 
$2\sigma$ error bars and weighting matrix $\widehat{\textbf{C}}_{\rm eff}$), the dashed black line represents the post-signal loss $2\sigma$ upper limit on the EoR signal, the dashed blue line represents the $2\sigma$ uniform-weighted power 
spectrum limit, and the solid green line denotes our $2\sigma$ theoretical noise prediction as calculated by Equation 
\eqref{eq:sense}. All three show good agreement, validating our analytical thermal noise calculation.

>>>>

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss3.pdf}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss4.pdf}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- 
foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of foregrounds peaks at a $k$-mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare empirically estimated inverse covariance weighting 
where $\textbf{C}$ is derived from the data (left), and projecting out the first eigenmode only (right). In the former 
case, signal loss arises from the coupling of the eigenmodes of $\widehat{\textbf{C}}$ to the data. For an empirically estimated $\widehat{\textbf{C}}$, its eigenvalues differ from those of the true covariance, where the coupling between EoR-dominated eigenmodes and the data can lead to loss.
Hence, there is negligible signal loss when assigning identical weights to all eigenmodes except the first, since we are only down-weighting the FG-dominating mode in this case.}
	\label{fig:toy_sigloss3}
\end{figure*}

@@

Because $\widehat{\textbf{C}}$ is estimated from the data, its eigenvectors and eigenvalues are strongly coupled to the particular data realization that was used to compute it. For example, the highest-valued eigenmode of $\widehat{\textbf{C}}$ (highest eigenvalue) describes the sinusoid foreground mode in the toy model (the peak in Figure \ref{fig:toy_sigloss3}). In Figure \ref{fig:toy_sigloss12} we show the estimated covariances of our toy model datasets along with $\widehat{\textbf{C}}^{-1}$ weighted data. The foreground sinusoid is clearly visible in $\widehat{\textbf{C}}_{\rm FG}$.

In general, the highest-valued eigenmodes of $\widehat{\textbf{C}}$ typically (but not always) describe bright foregrounds --- the most prominent shapes in a dataset. For these FG-dominating modes, where foregrounds outshine the EoR signal, down-weighting is beneficial. In some sense, we desire signal loss in this regime, if by `signal' we mean `foregrounds.' In this case it is beneficial for these eigenmodes to be coupled to the data in a way that produces loss. For our toy model, the successful suppression of the foreground mode is demonstrated in Figure \ref{fig:toy_sigloss3} by the missing foreground peak in the weighted power spectrum estimate (green).

@@

The danger of an empirically estimated covariance matrix comes mostly from not being able to describe the EoR-dominating eigenmodes of $
\textbf{C}$ accurately, for which the EoR signal is brighter than foregrounds. In such a case, the coupling between these modes to the data realization leads to the overfitting and subtraction of the EoR signal. More specifically, the coupling between the estimated covariance and the data is anti-correlated in nature (which is explained in more detail in Section \ref{sec:siglossmethod}), which leads to loss. Mis-estimating $\textbf{C}$ for EoR-dominated eigenmodes is therefore more harmful than for FG-dominated modes, and since the lowest-valued eigenmodes of an eigenspectrum are typically EoR-dominated, using this part of the spectrum for weighting is most dangerous.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield minimal 
signal loss. Recall that our toy model foreground is a sinusoid, so it can be perfectly described by a single eigenmode. Using the full dataset's (foreground plus EoR signal) empirical covariance, we can 
project out the first eigenmode and then flatten the rest of the spectrum to have eigenvalues of 1, thereby down-weighting the foreground-dominated mode more than the rest of the modes. Hence, we are changing the EoR-dominating part of the spectrum to be less coupled to the data, limiting the amount of over-fitting that can happen for those modes (i.e. only allowing over-fitting to occur for the first mode).

@@

We have shown how signal loss can arise due to the coupling of EoR-dominated eigenmodes to the data. We will next show how this effect is exacerbated by reducing the total number of independent 
samples in a dataset.

@@

The resulting eigenspectrum as compared to the green curve (FG + EoR) in Figure \ref{fig:toy_sigloss2} is shown in Figure 
\ref{fig:toy_sigloss15}. With 
fringe-rate filtering resulting in fewer independent modes, it becomes more difficult for the empirical covariance to estimate the 
true covariance matrix of the fringe-rate filtered data. This can be quantified by evaluating a convergence metric $\varepsilon(\Chat)$ for the empirical covariance, which we define as
\begin{equation}
\label{eq:converge}
\varepsilon (\Chat) \equiv \sqrt{\frac{\sum_{ij} (\widehat{C}_{ij} - {C}_{ij})^2}{\sum_{ij} {C}_{ij}^2}},
\end{equation}
where $\C$ is the true covariance matrix. To compute this metric, we draw different numbers of realizations (different draws of Gaussian noise) of our toy model EoR measurement, $\textbf{x}_{\rm EoR}$, and take their ensemble average. We then compare this to the ``true' covariance, which in our simulation is set to be the empirical covariance after a large number ($500$) of realizations. As shown in Figure \ref{fig:toy_sigloss16}, we perform this computation for a range of total independent ensemble realizations (horizontal axis) and the number of independent samples in the data following fringe-rate filtering (different colors). With more independent time samples (i.e. more realizations) in the data, one converges to the true fringe-rate filtered covariance more quickly.

@@

Just as important as the eigenvalues are the eigenvectors of our empirical covariances. In general, different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix whose true form is a diagonal matrix with eigenvalues spanning four orders of magnitude. We use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$ for the empirical eigenvectors $\widehat{\textbf{v}}$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves  in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. One sees that the highest-valued eigenmodes converge to their true eigenvectors more quickly. With only a small number of realizations, these empirically estimated modes already retain little correlation with the specific realizations of data that were used to form the empirical covariance. As we will see in the next section, using only these eigenmodes, which are less coupled to data realizations, minimizes signal loss. In contrast, the lower-valued eigenmodes retain more memory of the data realizations, which leads to correlations that induce signal loss. Said differently, a steep covariance eigenspectrum can be especially dangerous because the eigenmodes that converge the slowest are typically also EoR-dominated and are therefore susceptible to the most loss.  %\acl{Think about whether we need to do the Fourier space version as well.}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.31\textwidth]{plots/toy_sigloss15.pdf}
	\caption{Eigenspectrum of $\widehat{\textbf{C}}_{\rm FG+EoR}$, in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green). The dashed, steep eigenspectrum has a greater risk of signal loss because its lowest-valued eigenmodes are both more strongly coupled to the data (than those of the solid eigenspectrum) and, in this example, are also EoR-dominated.}
	\label{fig:toy_sigloss15}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/toy_sigloss17.pdf}
	\caption{The convergence level, as defined in Equation \eqref{eq:converge_eig}, of empirically estimated eigenvectors for different numbers of mock data realizations. The colors span from the 0th eigenmode (has the highest eigenvalue) to the 19th eigenmode (has the lowest eigenvalue), where they are ordered by eigenvalue in descending order. This figure shows that the highest-valued modes converge more quickly than the others, implying that weighting by the lowest-valued empirically estimated modes poses the most risk for loss. Since these modes are also typically EoR-dominated, this can lead to cosmological signal loss.}
	\label{fig:toy_sigloss17}
\end{figure}

@@

The third panel (bottom left) in Figure \ref{fig:toy_sigloss8} minimizes signal loss a different way --- 
by only using the first three eigenmodes of the estimated covariance. Recalling that our toy model foregrounds can be described entirely by the first eigenmode, this 
method intentionally projects out the highest-valued modes only by replacing all but the three highest weights in the 
eigenspectrum with $1$'s (equal weights). Again, avoiding the over-fitting of EoR-dominating modes which are coupled to the data results in negligible signal loss. However, we do 
not perfectly recover the shape of the EoR power spectrum because we lost information when ignoring the relative weights of most modes. While this case is illuminating for the toy model, in practice it is not obvious which eigenmodes are FG or EoR dominated (and they could be mixed as well), so determining which subset of modes to down-weight is not trivial.

The last regularization scheme we are highlighting here is setting $\widehat{\textbf{C}} \equiv \widehat{\textbf{C}} \circ \textbf{I}$ (element-wise multiplication), or inverse variance weighting (keeping only the diagonal elements of $\widehat{\textbf{C}}$). In the bottom right 
panel of Figure \ref{fig:toy_sigloss8}, we see that this method does not down-weight the foregrounds --- this regularization altered $\widehat{\textbf{C}}$ in a way where it is no longer coupled to any of the empirically estimated eigenmodes. For this toy model, 
our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by 
the identity, element-wise, results in a diagonal matrix, meaning we do not have any correlation information. Because of this, we do a poor job 
suppressing the foreground. But because we de-coupled the whole eigenspectrum from the data, we also avoid signal loss. Although this method did not successfully recover the EoR signal for this particular simulation, it is important that we show that there 
are many options for estimating a covariance matrix, and some may down-weight certain eigenmodes more than others based on the spectral nature 
of the components in a dataset. 
%One may imagine a situation where a particular systematic is contained to an isolated 
%frequency (such as radio frequency interference or crosstalk). In such a case, preserving only the diagonal elements of $
%\widehat{\textbf{C}}$ would be an effective way of removing this contamination.

@@

{\bf Small injection:}
%In this regime, the first term in Equation \eqref{eq:crossterm} is very similar to the power spectrum of data alone, namely $\text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \sim \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}$. They differ only in their covariance matrices, and when the injected signal is very small compared to the data, their difference can be neglected. 
%This is shown by the black and gray curves converging in both panels of Figure \ref{fig:sigloss_terms}. 
In this regime, the cross-terms (red) behave as noise averaged over a finite number of samples. Output values are Gaussian distributed around zero, spanning a range of values set by the injection level. This is because $\widehat{\textbf{R}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below). In fact, for the uniformly weighted case, the cross-term  $\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}$ is well modeled as a symmetric distribution with zero mean and width $\sqrt{P_e}\sqrt{P_x}$. We also note that in this regime, $\widehat{\textbf{P}}_{r}$ (black) approaches the data-only power spectrum value (gray) as expected. %In this limit we might imagine that the cross-correlation should also be small, enabling us to recover just the EoR term. However, examining the cross-term plotted in Figure \ref{fig:sigloss_terms} (red) in the identity case (right plot), we see that this term is still large enough to have a significant impact on $P_{\rm out}$. It is in this regime where we're dominated by biases from the cross-terms, causing the difference between the signal-only term (green) and $P_{\rm out}$ (black). We also note that the cross-terms behave as noise at small injections even in the inverse covariance weighted case (left plot), spanning a range of values set by the injection level. This is because $\widehat{\textbf{C}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below).

@@

{\bf In between:}
When the injected signal is of a similar amplitude to the data by itself, the situation becomes less straightforward. We see that 
the weighted injected power spectrum component mirrors the input power indicating little loss (i.e. the green curve follows the dotted black line), eventually 
departing from unity when the injected amplitude is well above the level of the data power spectrum. However, 
in this regime the cross-term has nearly the same amplitude, but with a negative sign. As explained below, this negativity is the result of cross-correlating inverse covariance weighted terms.  This negative component drives down the $P_{\rm out}$ estimator (black). We note that in \citetalias{ali_et_al2015}, signal loss was computed by only looking at the second term in Equation \eqref{eq:crossterm} (green), which incorrectly implies no loss at the data-only power spectrum level. Ignoring the effect of the negative power from the cross-terms is the main reason for under-estimating power spectrum limits in \citetalias{ali_et_al2015}.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.pdf}
	\caption{Illustration of the power spectrum amplitude of various power spectrum terms as a function of injected EoR power level summed into the 
data. Left: The empirically estimated inverse covariance weighted case used in \citetalias{ali_et_al2015}. The output power spectrum ($P_{\rm out}=\widehat{\textbf{P}}_r$) is shown in black, and the individual terms in Equation 
\eqref{eq:crossterm} are shown in blue, red, and green. The dotted diagonal black line indicates perfect 1:1 input-to-output mapping (no signal loss). The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we fit smooth polynomials to our data points to make an illustrative example.  The gray 
horizontal line is the power spectrum value of data alone, $\widehat{\textbf{P}}_{x}$ (it does not depend on injected power). The green signal-signal 
component is the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than $\widehat{\textbf{P}}_{r}$ (black) when the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlations between injection and data are not dominant (small and large $P_{\rm in}$), the cross-terms have a noise-like 
term with width $\sqrt{P_e}\sqrt{P_x}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to the couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data-only power spectrum value) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss). Right: The same power spectrum terms illustrated for the uniform weighted case.}
	\label{fig:sigloss_terms}
\end{figure*}

@@

\noindent where we can neglect the first term because $\textbf{x}$ is small (i.e. the large negative cross-term power in the left panel of Figure \ref{fig:sigloss_terms} occurs when the injected amplitude surpasses the level of the data-only power spectrum).  Without loss of generality, we will assume
an eigenbasis of $\textbf{e}$, so that $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$ is diagonal. The middle 
two terms, however, can have power in their off-diagonal terms due to the fact that, when averaging over a finite
ensemble, $\langle\textbf{xe}^\dagger\rangle_t$ is not zero.  As shown in Appendix C of \citet{parsons_et_al2014}%\footnote{This same appendix also gives a now-prescient warning about signal loss and the dangers of noise and sample variance in inverse covariance.}
, to leading order the inversion of a diagonal-dominant matrix like $\widehat{\textbf{C}}_{r}$ (from $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$) with smaller
off-diagonal terms results in a new diagonal-dominant matrix with negative off-diagonal terms. These off-diagonal
terms depend on both $\textbf{x}$ and $\textbf{e}$. Then, when $\widehat{\textbf{C}}^{-1}_{r}$ is multiplied into $\textbf{x}$,
the result is a vector that is similar to $\textbf{x}$ but
contains a residual correlation to $\textbf{e}$ from the off-diagonal components of $\widehat{\textbf{C}}^{-1}_{r}$. The
correlation is negative because the product $\widehat{\textbf{C}}_r^{-1}\textbf{x}$ effectively squares the $\textbf{x}$-dependence
of the off-diagonal terms in $\widehat{\textbf{C}}^{-1}_{r}$ while retaining the negative sign that arose from the inversion
of a diagonal-dominant matrix.
% XXX [ARP: we may want to write this all out explicitly in an appendix]

@@

The implementation details of the injection process require some more detailed explanation. In our code, we add a new realization of EoR to each independent bootstrap of data (see Section \ref{sec:Boot} for a description of PAPER's bootstrapping routine) with the goal of simultaneously capturing variance due to noise and signal loss. To limit computing time we perform $20$ realizations of each $P_{\rm in}$ level. We also run $50$ total EoR injection levels, yielding $P_{\rm in}$ values that range from $\sim$$10^{5}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$ to $\sim$10$^{11}$\,mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$, resulting in a total of $1000$ data points on our $P_{\rm in}$ vs. $P_{\rm out}$ grid.

@@

Power spectrum results for the noise simulation, which uses our full power spectrum pipeline, are shown in Figure 
\ref{fig:ps_noise}, where the black and gray points represent pre-signal loss estimated positive and negative power spectrum values, respectively (with 
$2\sigma$ error bars and weighting matrix $\widehat{\textbf{C}}_{\rm eff}$), the dashed black line represents the post-signal loss $2\sigma$ upper limit on the EoR signal, the dashed blue line represents the $2\sigma$ uniform-weighted power 
spectrum limit, and the solid green line denotes our $2\sigma$ theoretical noise prediction as calculated by Equation 
\eqref{eq:sense}. All three show good agreement to within a factor of two, which we believe is reasonable given the approximations we make in computing the analytic expression for thermal noise.
